---
title: "Sample experiment using Sliding Window Regression"
author: "Stefan Schrunner"
date: '2023-05-01'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Sample Experiment}
  %\VignetteEncoding{UTF-8}
---

# Introduction

<tt>SlidingWindowReg</tt> implements a regression model for hydrological time series data. In particular, multiple time-lagged windows with Gaussian kernel shape are estimated from training data. Then, each extracted feature is mapped to a target time series via multiple linear regression.

In this vignette, a sample experiment is performed to demonstrate that the model is able to identify model parameters correctly in a user-simulated setup.

## Data preparation \& experimental setup

First, we load the package and the sample watershed data:

```{r loading, message = FALSE}
library(SlidingWindowReg)

data("sampleWatershed")
```

As a second step, we divide the dataset into (a) a train set containing 30 years (75\% of the time series) and (b) a test set comprising the residual 10 years (25\% of the time series)

```{r splits}
train_inds <- 1 : (nrow(sampleWatershed) * 0.75)
```

## Run model

Given the data splits, we train the SlidingWindowReg model on the train set. For this purpose, the following parameters are used:

* <tt>iter</tt> number of iterations (maximum number of windows)
* <tt>runs</tt> number of independent model runs
* <tt>param\_selection</tt> method to determine the number of windows

```{r execute}
mod <- SlidingWindowReg::train(sampleWatershed$rain[train_inds],
                               sampleWatershed$gauge[train_inds],
                               iter = 3,
                               runs = 1,
                               parallel = FALSE,
                               param_selection = "best_bic")[[1]]
```

# Evaluation

In a first evaluation, we investigate the performance of the model on the test set. For this purpose, we compute the root mean squared error (RMSE):

```{r evaluate_rmse}
pred_on_test <- SlidingWindowReg::predict(sampleWatershed$rain[-train_inds],
                                          mix = mod$mix,
                                          param = mod$param)

print(rmse(pred_on_test, sampleWatershed$gauge[-train_inds]))
```

As a next step, we generate a plot demonstrating (a) the input time series, (b) the predicted output time series, and (c) the ground truth output.

```{r evaluate_plot, width = 20, height = 12}
plot_prediction(pred_on_test,
                sampleWatershed$gauge[-train_inds],
                sampleWatershed$rain[-train_inds])
```

## Model parameters

Finally, we can print the two main model parameters, consisting of:

* a matrix characterizing the location and size of the windows (column 1 indicates all center points on the time axis, column 2 indicates the variance)
* a vector of regression coefficients

```{r evaluate_param}
# window parameters
print(mod$param)

# regression parameters
print(mod$mix)
```

In order to visualize the estimated windows, a kernel plot is available in <tt>SlidingWindowReg</tt>:

```{r evaluate_param_plot, warning = FALSE, width = 12, height = 8}
plot_kernel(param = mod$param, 
            mix = mod$mix,
            include_text = FALSE)
```
